On the presentation of orthogonal groups over polynomial rings

### 2019-05-06 dl-paper.tex
Modified definition 8

\subsection{The subgroups $P_\alpha(0)$, $P_\alpha(*)$ and their properties.} Throughout this subsection $\alpha$ denotes some fixed root of $\Phi$.
\begin{df} \label{defP0}
Denote by $P_\alpha(0)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by the images of the following elements of $\St(\Phi, A[X])$ under the
 map induced by the embedding $A[X] \subseteq A[X, X^{-1}]$:
\begin{enumerate}
 \item $z_{\beta}(Xf, \xi), \beta \in \Phi \text{ such that }\alpha + \beta \in \Phi\text{ or } \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, X\xi),\ \beta \in \Phi \text{ such that }\alpha - \beta \in \Phi \cup \{0\};$
 \item $z_{\beta}(f, \xi),\ \alpha \perp \beta;$
 \item $x_{-\alpha}(X^2f),$
\end{enumerate}
In the above formulae and below we assume that $f \in M[X]$ and $\xi \in A[X]$.
{\bf TODO:} the above definition is somewhat incomplete, $P_\alpha(0)$ should also contain certain symbols, e.\,g. those from~\cref{lem:Tulenbayev-formula} below. \end{df}

Denote by $P_\alpha(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_\alpha(0)$ and the elements $x_{-\alpha}(mX)$, $m \in M$. The following lemma shows that this subgroup is sufficiently large.
\begin{lemma} \label{Pstar-large} The subgroup $P_\alpha(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof} Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$. 
Invoking~\eqref{Z5} of~\cref{Zrels} with $s = Xf$, $\eta = 1$ we obtain that $z_{\pm\alpha}(Xf, \xi)$ also belong to $P_{\alpha}(*)$. Thus, we have shown that $P_\alpha(*)$ contains $\St(\Phi, A[X], M[X])$.

@@

\begin{lemma}\label{P0_normal} The subgroup $P_\alpha(0)$ is normal in $P_\alpha(*)$. In particular, there is an exact sequence, which is split by the map $m \mapsto x_{-\alpha}(mX)$:
\[\xymatrix{1 \ar[r] & P_\alpha(0) \ar[r] & P_\alpha(*) \ar[r] & M^+ \ar@<5pt>@{-->}[l] \ar[r] & 1}.\]
\end{lemma}
\begin{proof} We need to verify that the conjugate by $x_{-\alpha}(mX)$ to the generators of $P_\alpha(0)$ listed in~Definition~\ref{defP0} belongs to $P_\alpha(0)$.
The assertion is obvious for the generators of type (3) and (4).

Let us verify the assertion for the generators of type (1) and (2). In the case $\beta - \alpha \in \Phi$ we obtain from Lemma~\ref{Zrels}:
\begin{align} z_{\beta}(Xf, \xi) ^ {x_{-\alpha}(mX)} = & x_{-\alpha} (- mX^2f\xi) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mX^2f) \cdot z_{\beta}(Xf, \xi), \label{eq3-1} \\
  z_{\beta}(f, X\xi) ^ {x_{-\alpha}(mX)} = & x_{-\alpha} (- mX^2f\xi ) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mXf) \cdot z_{\beta}(f, X\xi), \nonumber \\
  z_{\alpha}(f, X\xi) ^ {x_{-\alpha}(mX)} = & z_{\alpha}(f, X(\xi + m)). \nonumber \end{align}
The expressions in the right hand side of~\eqref{eq3-1} and~\eqref{eq3-2} are products of generators of type 4, 1, 1 and 4, 1, 2, respectively.

Finally, in the case $\alpha + \beta \in \Phi$ (only possible for the generator of type (1)) we have:
\begin{equation} \label{eq3-2} z_{\beta}(Xf, \xi) ^ {x_{-\alpha}(mX)} = x_{-\alpha} (mX^2f\xi ) \cdot x_{-\alpha-\beta} (N_{\beta,\alpha}\cdot mX^2f\xi^2) \cdot z_{\beta}(Xf, \xi). \end{equation}
It is clear that the latter expression is a product of generators of type $4, 1, 1$.
\end{proof}

@@

\begin{lemma}\label{P0_conj} The subgroup $P_\alpha(0)$ is stable under conjugation by $x_\alpha(aX^{-1})$ for arbitrary $a\in A$. \end{lemma}
\begin{proof} We need to verify that the conjugate by $x_\alpha(aX^{-1})$ to every generator $z_\beta(f, \xi)$ from~\cref{defP0} lies in $P_\alpha(0)$. 
First of all, consider the case $\alpha + \beta \in \Phi$, which is only possible for the generator of type 1:
\begin{equation}\label{eq:Zconj1} z_{\beta}(Xf, \xi) ^ {x_{\alpha}(aX^{-1})} = x_{\alpha} (- afg) \cdot x_{\alpha+\beta} (N_{\beta, \alpha}\cdot af) \cdot z_{\beta}(Xf, \xi). \end{equation}
Clearly, the expression in the right hand side is a product of generators of type 2, 2, 1. 
Now consider the case $\alpha - \beta \in \Phi$, which is possible for the generators of type 1 and 2.
Computing the conjugates of these generators using~\cref{Zrels} we obtain that
\begin{align} z_{\beta}(Xf, \xi) ^ {x_{\alpha}(aX^{-1})} = &  x_{\alpha} (af\xi) \cdot x_{\alpha-\beta} (N_{\beta,-\alpha}\cdot af\xi^2) \cdot z_{\beta}(Xf, \xi), \label{eq:Zconj2} \\
z_{\beta}(f, X\xi) ^ {x_{\alpha}(aX^{-1})} = & x_{\alpha} (af\xi) \cdot x_{\alpha-\beta} (N_{\beta,-\alpha}\cdot aXf\xi^2) \cdot z_{\beta}(f, X\xi). \nonumber \end{align}
The expressions in the right-hand side are products of generators of type 2, 2, 1 and 2, 1, 2.

Let us verify the assertion for the generator of type 4.
Choose $\beta\in \Phi$ such that $\alpha+\beta \in \Phi$. 
Substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$ and $\alpha = -\alpha - \beta$ into~\eqref{Z5} of~\cref{Zrels} we obtain that
\begin{multline} \label{eq:Zconj3} z_{-\alpha}(X^2f, aX^{-1}) = x_{-\alpha-\beta}(\epsilon Xf) \cdot x_{-\beta}(-af) \cdot x_{\beta}(aX^2 f) \cdot x_{-\alpha}(X^2f) \cdot \\
 \cdot z_{-\alpha-\beta}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha+\beta}(-\epsilon a^2 Xf) \cdot x_{\alpha}(- a^2 f) \cdot z_{-\beta}(a f, -X), \end{multline}
where $\epsilon = N_{-\alpha-\beta,\beta}$. 
Clearly, the expression in the right-hand side is a product of generators of type 1, 2, 1, 1, 1, 1, 2, 2.

It remains to verify the assertion for the remaining generator $z_\alpha(f, X\xi)$ of type 2.
Again choose $\beta$ to be any root such that $\alpha+\beta \in \Phi$.
Substituting in~\eqref{Z5} of~\cref{Zrels} $s = -\epsilon f$, $\xi = -\epsilon \xi$, $\eta=X$ and expressing $z_\alpha(f, X\xi)$ through other terms we obtain that
\begin{multline} \nonumber z_\alpha(f, X\xi) = x_{\alpha+\beta}(\epsilon Xf) \cdot x_{\beta}(-X^2 f \xi) \cdot x_{-\beta}(f \xi) \cdot x_\alpha(f) \cdot \\ 
 \cdot z_{\alpha+\beta}(-\epsilon X f, -\epsilon \xi) \cdot z_{-\beta}(-f\xi, -X) \cdot x_{-\alpha-\beta}(-\epsilon X f \xi^2) \cdot x_{-\alpha}(-X^2 f\xi^2), \end{multline}
where $\epsilon = N_{\alpha,\beta}$.  
The factors in the right-hand side are generators of type 1, 1, 2, 2, 1, 2, 1, 4. 
It is clear, that conjugation by $x_{\alpha}(aX^{-1})$ acts trivially on the fourth factor $x_\alpha(f)$, 
while for the other factors we have already verified that their conjugates belong to $P_{\alpha}(0)$. \end{proof}

>>>>

\subsection{The subgroups $P_\alpha(0)$, $P_\alpha(*)$ and their properties.} Throughout this subsection $\alpha$ denotes some fixed root of $\Phi$.
\begin{df} \label{defP0}
Denote by $P_\alpha(0)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by the images of the following elements of $\St(\Phi, A[X])$ under the
 map induced by the embedding $A[X] \subseteq A[X, X^{-1}]$:
\begin{enumerate}
 \item $z_{\beta}(Xf, \xi), \beta \in \Phi \text{ such that }\alpha + \beta \in \Phi\text{ or } \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, X\xi),\ \beta \in \Phi \text{ such that }\alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, \xi),\ \alpha \perp \beta;$
 \item $x_{-\alpha}(X^2f);$
 \item $x_{\alpha}(f).$ 
 \item Dennis--Stein symbols $\langle a, m \rangle$, $a\in A$, $m\in M$.
\end{enumerate}
In the above formulae and for the rest of this subsection we assume $f \in M[X]$ and $\xi \in A[X]$.
\end{df}

Denote by $P_\alpha(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_\alpha(0)$ and the elements $x_{-\alpha}(Xm)$, $m \in M$. The following lemma shows that this subgroup is sufficiently large.
\begin{lemma} \label{Pstar-large} The subgroup $P_\alpha(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof} Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$. 
Invoking~\eqref{Z5} of~\cref{Zrels} with $s = Xf$, $\eta = 1$ we obtain that $z_{\pm\alpha}(Xf, \xi)$ also belong to $P_{\alpha}(*)$. Thus, we have shown that $P_\alpha(*)$ contains $\St(\Phi, A[X], M[X])$.

@@

\begin{lemma}\label{P0_normal} The subgroup $P_\alpha(0)$ is normal in $P_\alpha(*)$. In particular, there is an exact sequence, which is split by the map $m \mapsto x_{-\alpha}(mX)$:
\[\xymatrix{1 \ar[r] & P_\alpha(0) \ar[r] & P_\alpha(*) \ar[r] & M^+ \ar@<5pt>@{-->}[l] \ar[r] & 1}.\]
\end{lemma}
\begin{proof} 
We need to verify that the conjugate by $x_{-\alpha}(mX)$ to every generator $g$ of $P_\alpha(0)$ listed in~Definition~\ref{defP0} belongs to $P_\alpha(0)$.
The assertion is obvious for the generators of type 3, 4 and 6.

Suppose $g$ has type 1 or 2 and $\beta - \alpha \in \Phi$. By~\cref{Zrels} we obtain that
\begin{align} z_{\beta}(Xf, \xi) ^ {x_{-\alpha}(mX)} = & x_{-\alpha} (- mX^2f\xi) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mX^2f) \cdot z_{\beta}(Xf, \xi), \label{eq3-1} \\
  z_{\beta}(f, X\xi) ^ {x_{-\alpha}(mX)} = & x_{-\alpha} (- mX^2f\xi ) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mXf) \cdot z_{\beta}(f, X\xi), \nonumber \end{align}
The expressions in the right hand sides of~\eqref{eq3-1} and~\eqref{eq3-2} are products of generators of type 4, 1, 1 and 4, 1, 2, respectively.

Suppose $g$ has type 1 and  $\alpha + \beta \in \Phi$. By~\cref{Zrels} 
\begin{equation} \label{eq3-2} z_{\beta}(Xf, \xi) ^ {x_{-\alpha}(mX)} = x_{-\alpha} (mX^2f\xi ) \cdot x_{-\alpha-\beta} (N_{\beta,\alpha}\cdot mX^2f\xi^2) \cdot z_{\beta}(Xf, \xi). \end{equation}
and the latter expression is a product of generators of type $4, 1, 1$.

Finally, suppose $g$ has type $5$.
Substituting in~\eqref{Z5} of~\cref{Zrels} $s = -\epsilon f$, $\xi = -\epsilon m$, $\eta=X$ and expressing $z_\alpha(f, mX)$ through other terms we obtain that
\begin{multline} \nonumber z_\alpha(f, mX) = x_{\alpha+\beta}(\epsilon Xf) \cdot x_{\beta}(-mX^2 f) \cdot x_{-\beta}(mf) \cdot x_\alpha(f) \cdot \\ 
 \cdot z_{\alpha+\beta}(-\epsilon X f, -\epsilon m) \cdot z_{-\beta}(-mf, -X) \cdot x_{-\alpha-\beta}(-\epsilon m^2X f) \cdot x_{-\alpha}(-m^2X^2 f), \end{multline}
and the latter expression is a product of generators of type 1, 1, 2, 5, 1, 2, 1, 4.
\end{proof}

@@

\begin{lemma}\label{P0_conj} The subgroup $P_\alpha(0)$ is stable under conjugation by $x_\alpha(aX^{-1})$ for arbitrary $a\in A$. \end{lemma}
\begin{proof} We need to verify that the conjugate by $x_\alpha(aX^{-1})$ to every generator $g = z_\beta(f, \xi)$ from~\cref{defP0} lies in $P_\alpha(0)$. 
The assertion is clear for generators of type 3, 5 and 6.
First of all, consider the case $\alpha + \beta \in \Phi$, which is only possible for the generator of type 1:
\begin{equation}\label{eq:Zconj1} z_{\beta}(Xf, \xi) ^ {x_{\alpha}(aX^{-1})} = x_{\alpha} (- afg) \cdot x_{\alpha+\beta} (N_{\beta, \alpha}\cdot af) \cdot z_{\beta}(Xf, \xi). \end{equation}
Clearly, the expression in the right hand side is a product of generators of type 5, 2, 1.

Now consider the case $\alpha - \beta \in \Phi$, which is possible for the generators of type 1 and 2.
Computing the conjugates of these generators using~\cref{Zrels} we obtain that
\begin{align} z_{\beta}(Xf, \xi) ^ {x_{\alpha}(aX^{-1})} = &  x_{\alpha} (af\xi) \cdot x_{\alpha-\beta} (N_{\beta,-\alpha}\cdot af\xi^2) \cdot z_{\beta}(Xf, \xi), \label{eq:Zconj2} \\
z_{\beta}(f, X\xi) ^ {x_{\alpha}(aX^{-1})} = & x_{\alpha} (af\xi) \cdot x_{\alpha-\beta} (N_{\beta,-\alpha}\cdot aXf\xi^2) \cdot z_{\beta}(f, X\xi). \nonumber \end{align}
The expressions in the right-hand side are products of generators of type 5, 2, 1 and 5, 1, 2.

Let us verify the assertion for the generator of type 4.
Choose $\beta\in \Phi$ such that $\alpha+\beta \in \Phi$. 
Substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$ and $\alpha = -\alpha - \beta$ into~\eqref{Z5} of~\cref{Zrels} we obtain that
\begin{multline} \label{eq:Zconj3} z_{-\alpha}(X^2f, aX^{-1}) = x_{-\alpha-\beta}(\epsilon Xf) \cdot x_{-\beta}(-af) \cdot x_{\beta}(aX^2 f) \cdot x_{-\alpha}(X^2f) \cdot \\
 \cdot z_{-\alpha-\beta}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha+\beta}(-\epsilon a^2 Xf) \cdot x_{\alpha}(- a^2 f) \cdot z_{-\beta}(a f, -X), \end{multline}
where $\epsilon = N_{-\alpha-\beta,\beta}$. 
Clearly, the expression in the right-hand side is a product of generators of type 1, 2, 1, 4, 1, 1, 5, 2.

\begin{comment}
It remains to verify the assertion for the remaining generator $z_\alpha(f, X\xi)$ of type 2.
Again choose $\beta$ to be any root such that $\alpha+\beta \in \Phi$.
Substituting in~\eqref{Z5} of~\cref{Zrels} $s = -\epsilon f$, $\xi = -\epsilon \xi$, $\eta=X$ and expressing $z_\alpha(f, X\xi)$ through other terms we obtain that
\begin{multline} \nonumber z_\alpha(f, X\xi) = x_{\alpha+\beta}(\epsilon Xf) \cdot x_{\beta}(-X^2 f \xi) \cdot x_{-\beta}(f \xi) \cdot x_\alpha(f) \cdot \\ 
 \cdot z_{\alpha+\beta}(-\epsilon X f, -\epsilon \xi) \cdot z_{-\beta}(-f\xi, -X) \cdot x_{-\alpha-\beta}(-\epsilon X f \xi^2) \cdot x_{-\alpha}(-X^2 f\xi^2), \end{multline}
where $\epsilon = N_{\alpha,\beta}$.  
The factors in the right-hand side are generators of type 1, 1, 2, 2, 1, 2, 1, 4. 
It is clear, that conjugation by $x_{\alpha}(aX^{-1})$ acts trivially on the fourth factor $x_\alpha(f)$, 
while for the other factors we have already verified that their conjugates belong to $P_{\alpha}(0)$. 
\end{comment}
\end{proof}
